Learning constraints in spreadsheets and tabular data

### 2016-05-30 Paper/ECAI/tabular.tex
Several corrections

\newcommand{\sg}{B}
\newcommand{\sbl}[1]{\ensuremath{B_{\textit{#1}}}}

@@

The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.  It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
very few examples.  Our approach contrasts with these in that it focusses
on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional constraint satisfaction problems (CSPs). It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.

\samuel{Move this up a paragraph}
The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given. %, and discovers column- and row-wise constraints.
We demonstrate the utility of our approach in an experimental evaluation.
Moreover, we sketch additional application scenario's such as autocompletion and error detection.
\\\\
This paper is organized as follows.
Section \ref{sec:formalization} introduces concepts relevant to our approach. Section \ref{sec:approach} presents the problem statement and the approach. Section \ref{sec:evaluation} presents the evaluation of the approach. Section \ref{sec:applications} shows how our system can be used for applications. Section \ref{sec:related_work} presents a more detailed discussion of the related work. Section \ref{sec:conclusions} provides conclusions.

@@

  \begin{subfigure}{.65\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo2.png}
  \end{center}
  \vspace{-10pt}
  \caption{
    Example spreadsheet (black words and numbers only).
    Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
    Most tables only contain type-consistent columns, however, the data in table~$T_2$ can be used as both column- and row vectors.
    This example was created by combining several smaller Excel sheets originating from an exercise session.
    \samuel{A: I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
  }
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.35\textwidth}
  {\footnotesize
    \begin{align*}
      % Table 1
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
%      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}\\
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^1, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^1 \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^1 , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^1 \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^1 \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
	  % Table 2
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
	  % Table 4
      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^2 \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
	  % Table 6
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
  $^1$(Spurious constraints) $^2$(Not present in original source)}
  \label{fig:sol_example}
\end{subfigure}
  \caption{Running example}
\end{figure*}

@@

We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
%
An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.

\begin{table*}
\caption{
  Overview of constraint templates implemented in \sname.
  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows the version for sum but \sname also supports max, min, average, product and count.
  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector.
  Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.  {\small\luc{there are some fancy constraints here. it might be good to say something about them ... like an example of one of the more complex ones in the text ... maybe state what conditional means, what aggregate ... }}
}
\label{table:constraints}
  {\centering
  \begin{tabularx}{\textwidth}{l X X}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
    $\ecalldiff{\sg_x}^*$
      & $\discrete(\sg_x)$

@@

      & $B_r[i] = \sum ($row~$i$ of $\mathbf{B_x})$
      \\ \hline

@@

      & $B_r[i] = \sum ($column~$i$ of $\mathbf{B_x})$
      \\ \hline

@@

@@

\item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
This can concern any property of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
\\\\
\samuel{A: I'm a bit confused by the fact that you seem to be mixing the formal way of writing these triples with textual explanation, both in the example in the text and in the table; I'd prefer to see the formal thing and then get the explanation, as that's what I'd need to use your approach}
\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
Given an assignment of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks. %holds, in practice this will be a function that can be called when
%the arguments are full instantiated to decide whether the constraint is satisfied or not.
In logical and relational learning this is known as the background knowledge. %, which provides the definition of the predicate.
\end{itemize}

\begin{example}
Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:

@@

\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
\newcommand{\defc}{\ensuremath{\format{Def}_s}}
%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming and logical relational learning, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
Under this interpretation, the term constraint and constraint template can be used interchangeably.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/tabletool.png}
  \caption{This figure shows a simple table selection tool. A user selects rectangular ranges of cells that correspond to (headerless) tables, an explicit orientation can also be provided.}
  \label{fig:visual_tool}
\end{figure}

@@

%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
\begin{example}
  Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
  An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
  Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition: $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$, where $B[i][j]$ indicates the $i$th value of the $j$th vector in $B$.
  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$. \luc{why [j][i] and not [j,i] ?}
\end{example}

@@

\begin{example}
  Consider $\ecfkey{B_{fk}}{B_{pk}}$ whose signature includes $\ecalldiff{B_{pk}}$ and assume that~$C$ are the constraints found for \textit{ALLDIFFERENT}.
  Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed. \samuel{is it clear now?}
\end{example}

@@

A last type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
For example, consider the constraint \ecsumc{B_r}{B[1:n]} where $B[k:n]$ consists of only zeros, then \ecsumc{B_r}{B[1:j]} will be true for all $k - 1 \leq j \leq n$.
In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.

As a result, in some cases a maximal subblock might falsely include irrelevant columns.
Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.

Alternatively, we can output all valid argument assignments instead of only maximal subblocks, e.g. $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.

@@

\subsubsection*{Q1. How accurate \sname?}
Our implementation achieves a recall of~$0.97$ for relevant constraints, i.e. it finds~90 of~93 relevant constraints in the benchmark suite.
The recall for supported constraints is $1.0$ (90 of 90).

@@

\subsubsection*{Q3. How fast is the algorithm?}
Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
For the \benchsize~spreadsheets in the benchmark our implementation ran in ${\runtotal}s \pm {\runtotalstd}s$.
The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).

\begin{table}
  \centering
  \begin{tabular}{lll}
    & \textbf{Total} & \textbf{Average per spreadsheet} \\
    \textbf{Constraints} & $93$ & $2.91$ \\
    \textbf{Accuracy} & $96.77\%$ & $94.27\%$ \\
    \textbf{Redundant} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
    \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
  \end{tabular}
  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
\end{table}
\samuel{Expand summary}
\sergey{would it make sense to specify the size of a spreadsheet in the number of non-zero cells?}

\samuel{Single paragraph / section}
\paragraph*{Dependencies}
In order for the algorithm to run efficiently it is crucial to use dependencies and find constraints incrementally whenever possible.
This avoids some of the explosion of combinations for constraints that have many arguments.
For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to below 1 second.
Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.

>>>>

\newcommand{\sg}{B}
\newcommand{\sbl}[1]{\ensuremath{B_{\textit{#1}}}}
\newcommand{\bsbl}[1]{\ensuremath{\mathbf{B_{\textit{#1}}}}}

@@

% Moved up
The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given. %, and discovers column- and row-wise constraints.
We demonstrate the utility of our approach in an experimental evaluation.
Moreover, we sketch additional application scenario's such as autocompletion and error detection.

The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.  It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
very few examples.  Our approach contrasts with these in that it focusses
on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional constraint satisfaction problems (CSPs). It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
\\\\
This paper is organized as follows.
Section \ref{sec:formalization} introduces concepts relevant to our approach. Section \ref{sec:approach} presents the problem statement and the approach. Section \ref{sec:evaluation} presents the evaluation of the approach. Section \ref{sec:applications} shows how our system can be used for applications. Section \ref{sec:related_work} presents a more detailed discussion of the related work. Section \ref{sec:conclusions} provides conclusions.

@@

  \begin{subfigure}{.65\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo2.png}
  \end{center}
  \vspace{-10pt}
  \caption{
    Example spreadsheet (black words and numbers only).
    Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
    Most tables only contain type-consistent columns, however, the data in table~$T_2$ can be used as both column- and row vectors.
    This example was created by combining several smaller Excel sheets originating from an exercise session.
    % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
  }
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.35\textwidth}
  {\footnotesize
    \begin{align*}
      % Table 1
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
%      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}\\
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^1, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^1 \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^1 , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^1 \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^1 \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
	  % Table 2
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
	  % Table 4
      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^2 \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
	  % Table 6
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
  $^1$(Spurious constraints) $^2$(Not present in original source)}
  \label{fig:sol_example}
\end{subfigure}
  \caption{Running example}
\end{figure*}

@@

We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
%
An example of block containment is $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.

\begin{table*}
\caption{
  Overview of constraint templates implemented in \sname.
  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows the version for sum but \sname also supports max, min, average, product and count.
  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector.
  Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
  {\small\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
}
\label{table:constraints}
  {\centering
  \begin{tabularx}{\textwidth}{l X X}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
    $\ecalldiff{\sg_x}^*$
      & $\discrete(\sg_x)$

@@

      & $B_r[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$
      \\ \hline

@@

      & $B_r[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \mathit{column}(i, \bsbl{x})[j]$
      \\ \hline

@@

\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
\newcommand{\defc}{\ensuremath{\format{Def}_s}}

@@

\item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
This can concern any property of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
We can capture this bias for a template~$s$ using a predicate $\sigc$.
\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
Given an assignment of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks. %holds, in practice this will be a function that can be called when
%the arguments are full instantiated to decide whether the constraint is satisfied or not.
In logical and relational learning this is known as the background knowledge.
We introduce the predicate \defc~to capture this background knowledge for a template~$s$.
\end{itemize}

\begin{example}
The constraint templates implemented in \sname are defined in Table~\ref{table:constraints}.
A non-trivial example is, for example, the constraint template for the row-based sum:

@@

With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming and logical relational learning, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
Under this interpretation, the term constraint and constraint template can be used interchangeably.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/tabletool.png}
  \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding (headerless) tables, an explicit orientation can also be provided.}
  \label{fig:visual_tool}
\end{figure}

@@

%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
\begin{example}
  Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
  An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
  Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition:
  $\forall i, \sbl{r}[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
  In Figure~\ref{fig:main_example} there is exactly one such subassignment: $(\range{T_1}{\rangeall}{7}, \range{T_1}{\rangeall}{\rangeto{3}{6}})$.
\end{example}

@@

\begin{example}
  Consider $\ecfkey{B_{fk}}{B_{pk}}$ whose signature includes $\ecalldiff{B_{pk}}$ and assume that~$C$ are the constraints found for \textit{ALLDIFFERENT}.
  Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed.
\end{example}

@@

A last type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
For example, consider the constraint \ecsumc{B_r}{B[1{:}n]} where $B[k{:}n]$ consists of only zeros, then \ecsumc{B_r}{B[1{:}j]} will be true for all $k - 1 \leq j \leq n$.
In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.

As a result, in some cases a maximal subblock might falsely include irrelevant columns.
Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$.
Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.

Alternatively, we can output all valid argument assignments instead of only maximal subblocks, e.g. $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.

@@

\subsubsection*{Q1. How accurate is \sname?}
Our implementation achieves a recall of~$0.97$ for relevant constraints, i.e. it finds~90 of~93 relevant constraints in the benchmark suite.
The recall for supported constraints is $1.0$ (90 of 90).

@@

\subsubsection*{Q3. How fast is \sname?}
Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
For the \benchsize~spreadsheets in the benchmark our implementation ran in ${\runtotal}s \pm {\runtotalstd}s$.
The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).

\begin{table}
  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
  \centering
  \begin{tabular}{lll}
    & \textbf{Total} & \textbf{Average per spreadsheet} \\
    \textbf{Relevant constraints} & $93$ & $2.91$ \\
    \textbf{Recall} & $96.77\%$ & $94.27\%$ \\
    \textbf{Precision} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
    \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
  \end{tabular}
\end{table}
\samuel{Expand summary}
\sergey{would it make sense to specify the size of a spreadsheet in the number of non-zero cells?}

We explained various optimizations, most notably the exploitation of dependencies.
Using the dependencies to find constraints incrementally cuts down large parts of the search trees, especially for constraints that have many arguments.
Essentially, one obtains a more specialized search without much effort.

In order for the algorithm to run efficiently it is crucial to use dependencies and find constraints incrementally whenever possible.
This avoids some of the explosion of combinations for constraints that have many arguments.
For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to below 1 second.
Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
